Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements

### 2018-06-09 Conference/AAS 2015 OD EKF/AAS15_.tex
added Conference folder

%\documentclass[letterpaper, preprint, paper,11pt]{AAS}	% for preprint proceedings
%\documentclass[letterpaper, paper,11pt]{AAS}		% for final proceedings (20-page limit)
%\documentclass[letterpaper, paper,12pt]{AAS}		% for final proceedings (20-page limit)
\documentclass[letterpaper, paper,10pt]{AAS}		% for final proceedings (20-page limit)
%\documentclass[letterpaper, submit]{AAS}			% to submit to JAS

\usepackage{bm}
\usepackage{amssymb,amsmath,amsthm,times,graphicx,subfigure,tabularx,booktabs,colortbl,multirow,threeparttable}
\usepackage{subfigure}
%\usepackage[notref,notcite]{showkeys}  % use this to temporarily show labels
\usepackage[colorlinks=true, pdfstartview=FitV, linkcolor=black, citecolor= black, urlcolor= black]{hyperref}
\usepackage{overcite}
\usepackage{footnpag}			      	% make footnote symbols restart on each page

\usepackage{color}
\newcommand{\hilight}[1]{\colorbox{green}{#1}}

\newtheorem{prop}{Proposition}

\newcommand{\abs}{\mathop{\mathrm{abs}}\nolimits}
\newcommand{\diag}{\mathop{\mathrm{diag}}\nolimits}
\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\ave}[1]{\mathrm{E}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathfrak{d}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}

@@

\PaperNumber{23-451}

@@

\author{Evan Kaufman\thanks{Doctoral Student, Mechanical and Aerospace Engineering, George Washington University, 801 22nd St NW, Washington, DC 20052, Email: evankaufman@gwu.edu.},  
T. Alan Lovell\thanks{Research Aerospace Engineer, Air Force Research Laboratory, Space Vehicles Directorate, Kirtland AFB, NM.},
\ and Taeyoung Lee\thanks{Assistant Professor, Mechanical and Aerospace Engineering, George Washington University, 801 22nd St NW, Washington, DC 20052, Tel: 202-994-8710, Email: tylee@gwu.edu.}
\thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}
}

@@

\begin{abstract}
%An angles-only extended Kalman filter (EKF), based off of two-body motion, is developed to estimate the position of a Deputy satellite relative to a chief satellite.
A new nonlinear observability measure is proposed for relative orbit determination when lines-of-sight between satellites are measured only. It corresponds to a generalization of the observability Gramian in linear dynamic systems to the nonlinear relative orbit dynamics represented by the two-body problems. An extended Kalman filter (EKF) is adapted to this problem and is evaluated with various gravitational harmonics and initial orbital determination (IOD) predictions. Extensive results illustrate correspondence between the proposed observability measure with filtering errors. An extensive numerical analysis in realistic scenarios includes satellite propagation of the two-body problem the $J_2$ perturbation effects.
\end{abstract}

@@

Space-based surveillance or relative navigation is desirable for many spacecraft missions, such as formation control and rendezvous. Spacecraft maneuvers based only on on-board measurements reduce the total operating cost significantly, and it improves safety against communication interruptions with ground stations.  Relative navigation between spacecraft in close-proximity essentially corresponds to space-based orbit determination.  In particular, vision-based navigation and estimation of relative orbits have received attention recently, since optical sensors have the desirable properties of low cost and minimal maintenance, while providing accurate line-of-sight measurements, or angles from a chief to a deputy.

@@

However, this result does not provide any information of the degree of observability. The main contribution of this paper is proposing a new  measure of observability for relative orbit determination with angles-only measurements. For a given initial condition of relative orbits, the proposed observability measure determines how much the corresponding relative orbits are easy or difficult to estimate, thereby providing a quantitative measure for relative orbit determination.

@@

\section{Nonlinear Relative Orbital Dynamics}\label{sec:ND}

Consider two satellites orbiting around the Earth, where each satellite is modeled as a rigid body. Suppose that a \textit{chief} satellite is on a circular orbit with a pre-determined orbital radius of $a\in\Re$.

Define a local-vertical, local-horizontal (LVLH) frame as follows. Its origin is located at the chief satellite. The $x$-axis is along the radial direction from the Earth to the chief, and the $y$-axis is along the velocity vector of the chief. The $z$-axis is normal to the orbital plane, and it is parallel to the angular momentum vector of the chief. The LVLH frame is rotating with the angular velocity of $\mathbf{\omega}=[0,0,n]^T\in\Re^3$, where $n=\sqrt{\frac{\mu}{a^3}}$ is the mean motion of the chief satellite, and $\mu$ denotes the gravitational parameters of the Earth. Note that the velocity of the chief is given by $\mathbf{v}_{chief}=[0,na,0]^T\in\Re^3$. Let the relative position of a \textit{deputy} satellite with respect to the chief satellite be given by $\mathbf{r}=[x,y,z]^T\in\Re^3$ in the LVLH frame.

\subsection{Nonlinear Equations of Motion}

Nonlinear equations of motion for the relative motion of the deputy with respect to the chief can be derived as follows based on Lagrangian mechanics~\cite{LovLeePISSFD14}.

\paragraph{Lagrangian}

Considering that the LVLH frame is rotating with the angular velocity of $\mathbf{\omega}=[0,0,n]^T$, the inertial velocity of the deputy satellite is given by
\begin{align*}
\mathbf{v}_{deputy}
=
\mathbf{v}_{chief} + \dot{\r} + \mathbf{\omega}\times \mathbf{r} = 
\begin{bmatrix}
0 \\ na \\ 0
\end{bmatrix}
+
\begin{bmatrix}
\dot x \\ \dot y \\ \dot z
\end{bmatrix}
+
\begin{bmatrix}
-n y \\ nx \\ 0
\end{bmatrix}
=
\begin{bmatrix}
\dot x - ny \\ \dot y + nx + na \\ \dot z
\end{bmatrix}.
\end{align*}
Therefore, the (normalized) kinetic energy of the deputy satellite is 
\begin{align*}
T = \frac{1}{2}\|\mathbf{v}_{deputy}\|^2 = \frac{1}{2} \braces{(\dot x -ny)^2+(\dot y + nx +na)^2 + \dot z^2}.
\end{align*}
The location of the Earth from the chief is given by $[-a,0,0]^T$ in the LVLH frame. Therefore, the position vector of the deputy from the center of the Earth is given by 
$\r_a=[x+a,y,z]^T\in\Re^3$. The gravitational potential energy is 
\begin{align*}
U = -\frac{\mu}{\sqrt{(x+a)^2 + y^2 + z^2}} = -\frac{\mu}{\|\r_a\|}.
\end{align*}
From the above equations, the Lagrangian of the deputy satellite is expressed in terms of $(x,y,z)$ as 
\begin{align}
L & = T-U = \frac{1}{2} \braces{(\dot x -ny)^2+(\dot y + nx +na)^2 + \dot z^2}
+\frac{\mu}{\sqrt{(x+a)^2 + y^2 + z^2}}.
\end{align}

\paragraph{Euler-Lagrange Equations}

Using the Euler-Lagrange equations, given by
\begin{align*}
\frac{d}{dt}\deriv{L}{\dot q}-\deriv{L}{q}=0,
\end{align*}
for $q\in\{x,y,z\}$, we obtain the nonlinear equations of motion for the relative orbit as follows.
\begin{align}
\ddot x - 2n\dot y -n^2 x&=n^2 a - \frac{\mu (x+a)}{((x+a)^2 + y^2 + z^2)^{3/2}},\label{eqn:ddotx}\\
\ddot y + 2n\dot x -n^2 y &=   - \frac{\mu y}{((x+a)^2 + y^2 + z^2)^{3/2}},\label{eqn:ddoty}\\
\ddot z &= - \frac{\mu z}{((x+a)^2 + y^2 + z^2)^{3/2}}.\label{eqn:ddotz}
\end{align}

These can be written as the standard form of the state equation,
\begin{align}
\dot{\mathbf{x}} = f (\mathbf{x}), \label{eqn:xxdot}
\end{align}
where the state vector is $\mathbf{x}=[x,y,z,\dot x, \dot y,\dot z]^T\in\Re^N$ with $N=6$, and
\begin{align}
f(\mathbf{x}) 
%= \begin{bmatrix}
%\dot x\\
%\dot y\\
%\dot z\\
%2n\dot y +n^2 (x+a) - \dfrac{\mu (x+a)}{((x+a)^2 + y^2 + z^2)^{3/2}}\\
%-2n\dot x +n^2 y  - \dfrac{\mu y}{((x+a)^2 + y^2 + z^2)^{3/2}}\\
%- \dfrac{\mu z}{((x+a)^2 + y^2 + z^2)^{3/2}}
%\end{bmatrix}
=\begin{bmatrix}
\dot \r \\
-2\omega\times \dot\r - \omega\times(\omega\times \r_a) -\dfrac{\mu \r_a}{\|\r_a\|^3}
\end{bmatrix}.\label{eqn:f}
\end{align}
%where $\r=[x,y,z]^T\in\Re^3$, $\r_a = [x+a,y,z]^T\in\Re^3$, and $\dot\r =[\dot x,\dot y,\dot x]^T\in\Re^3$.

\subsection{Line-of-sight Measurement}

We assume that the line-of-sight from the chief to the deputy is measured by onboard sensing, such as optical sensors. The measurement is represented by the unit-vector of the relative position vector, i.e.,
\begin{align}
\y = h(\mathrm{x}) =\frac{\r}{\|\r\|},\label{eqn:y}
\end{align}
where $\y\in\Re^M$ with $M=3$, and it satisfies $\|\y\|=1$.

\section{Nonlinear Observability Criteria}\label{sec:OC}

Based on the nonlinear dynamic model presented at the previous section, here we analyze the observability of the relative orbit with line-of-sight measurements.
This section serves as a summary of~\cite{LovLeePISSFD14}.

\subsection{Observability Criteria for Nonlinear Systems}

Observability of nonlinear systems has been studied in~\cite{HerKreITAC77}, and it is summarized as follows. For a given nonlinear dynamic system described with \refeqn{xxdot} and \refeqn{y}, a pair of points $\x_0$ and $\x_1$ are called \textit{indistinguishable} if the outputs of the corresponding solutions starting from each of $\x_0$ and $\x_1$ are identical for a certain time period. The systems is \textit{locally weakly observable} at $\x_0$, if there exists an open neighborhood $V$ of $\x_0$ such that for every open neighborhood $U$ of $\x_0$ contained in $V$, the only indistinguishable point to $\x_0$ is the point $\x_0$ itself.

The Lie-derivative of the output $h(\x)$ along $f(\x)$ is defined as follows:
\begin{align*}
L_f h(\x) = \deriv{h(\x)}{\x}f(\x)\in\Re^{M\times 1},
\end{align*}
which corresponds to the directional derivative of $h(\x)$ along $f(\x)$. For a non-negative integer $i$, the $i$-th order Lie-derivative is defined by induction as $L_f^i h = L_f (L_f^{i-1} h)$ with $L_f^0 h = h$. Define an observability matrix $\mathcal{O}\in\Re^{NM\times N}$ as
\begin{align*}
\mathcal{O}(\x_0) = \deriv{}{\x} \begin{bmatrix} L_f^0 h(\x) \\ L_f^1 h(\x)\\ \vdots \\L_f^{N-1} h(\x)\end{bmatrix}\bigg|_{\x=\x_0}.
\end{align*}
It has been shown that the system is locally weakly observable at $\x_0$ if the rank of the observability matrix $\mathcal{O}(\x_0) = N$. When applied to linear dynamics, this yields the well-known observability rank condition for linear systems. Note that when there is more than a single measurement, i.e., $M>1$, the observability rank condition can be satisfied without a need for computing the higher-order Lie derivatives up to the $N-1$-th order.

\subsection{Observability Criteria for Relative Orbital Dynamics}

\paragraph{Observability Matrix}
We apply the above observability criteria for the nonlinear relative orbital dynamics. Using the fact that $\x=[\r^T,\dot \r^T]$, the observability matrix of the relative orbital dynamics can be written as
\begin{align}
\mathcal{O} =
\begin{bmatrix}
\deriv{\y}{\r} & \deriv{\y}{\dot\r}\\
\deriv{\dot\y}{\r} & \deriv{\dot\y}{\dot\r}\\
\deriv{\ddot\y}{\r} & \deriv{\ddot\y}{\dot\r}
\end{bmatrix}
\triangleq
\begin{bmatrix}
\mathcal{O}_{00} & 0_{3\times 3}\\
\mathcal{O}_{10} & \mathcal{O}_{11}\\
\mathcal{O}_{20} & \mathcal{O}_{21}
\end{bmatrix}.\label{eqn:OO}
\end{align}
Here we consider the observability matrix obtained by up to the second order Lie derivatives of the measurement due to complexity. But, this still provides sufficient conditions for observability.

After straightforward but tedious algebraic manipulations using the following identity repeatedly,
\begin{align*}
\delta \parenth{\frac{1}{\|\mathbf{r}\|^i}} 
& = -i\frac{ \mathbf{r}^T \delta \mathbf{r}}{\|\mathbf{r}\|^{i+2}},
\end{align*}
for any positive integer $i$, we can show that each of the sub-matrices $\mathcal{O}_{ij}$ of the observability matrix $\mathcal{O}$ is given by
\begin{align}
\mathcal{O}_{00} & = \deriv{\y}{\r} = \frac{1}{\|\r\|}(I - \y\y^T),\label{eqn:O_00}\\
\mathcal{O}_{10} & = \deriv{\dot\y}{\r} = \deriv{}{\r} \parenth{\deriv{\y}{\r}\dot \r}= -\frac{1}{\|\r\|^2}\braces{
{\dot\r\y^T +\y^T\dot\r I +\y\dot\r^T}
- 3\y\y^T\dot\r \y^T},\\
\mathcal{O}_{11} & = \deriv{\dot\y}{\dot\r} = \deriv{}{\dot\r} \parenth{\deriv{\y}{\r}\dot \r}= \deriv{\y}{\r} = \mathcal{O}_{00},\label{eqn:O11}\\
\mathcal{O}_{20} & = \deriv{\ddot\y}{\r} = -\dfrac{2\dot\r \dot\r^T  +\dot\r^T\dot\r I}{\|\r\|^3}
+3\dfrac{2(\r^T\dot\r) \dot\r\r^T +(\dot\r^T\dot\r)\r\r^T
+(\r^T\dot\r)^2 I
+2(\r^T\dot\r)\r  \dot\r^T
}{\|\r\|^5}
- 15 \dfrac{(\r^T\dot\r)^2\r  \r^T}{\|\r\|^7}\nonumber \\
&\quad -\dfrac{\ddot{\r}\r^T +\r^T\ddot{\r} I +\r \ddot{\r}^T}{\|\r\|^3}
+ 3 \dfrac{(\r^T\ddot{\r})\r \r^T}{\|\r\|^5}
+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}},\\
\mathcal{O}_{21} & = \deriv{\ddot\y}{\dot\r} = \deriv{}{\dot\r}\parenth{\deriv{}{\r}\parenth{\deriv{\y}{\r}\dot\r}\dot\r + \deriv{\y}{\r}\ddot\r} = 2 \deriv{}{\r} \parenth{\deriv{\y}{\r}\dot\r} + \deriv{\y}{\r} \deriv{\ddot\r}{\dot\r}
\nonumber\\
& \quad = 2\mathcal{O}_{10} 
- 2 \mathcal{O}_{00}[\omega]_\times,\label{eqn:O_21}
\end{align}
where $[\omega]_\times\in\Re^{3\times 3}$ is defined as
\begin{align*}
[\omega]_\times = \begin{bmatrix} 0 & -n & 0 \\ n & 0 & 0 \\ 0 & 0 & 0 \end{bmatrix}.
\end{align*}
These expressions were fully verified by the Matlab symbolic computation toolbox.

After some algebraic manipulations, we can show that the sub-matrices satisfy the following identities. 
\begin{align}
\mathcal{O}_{10}\r & = 
%-\frac{1}{\|\r\|}\braces{{\dot\r\y^T +\y^T\dot\r I +\y\dot\r^T}- 3\y\y^T\dot\r \y^T}\y\\
%&=-\frac{1}{\|\r\|}\braces{\dot\r +(\y^T\dot\r)\y +\y(\dot\r^T\y)- 3\y(\y^T\dot\r) }\\
%&=-\frac{1}{\|\r\|}\braces{\dot\r -\y(\y^T\dot\r) }=
-\mathcal{O}_{00}\dot\r,\label{eqn:O10r}\\
\mathcal{O}_{10}\dot\r & = -\frac{1}{\|\r\|^2}\braces{
{2\dot\r(\y^T\dot \r)  +\y(\dot\r^T\dot \r)}
- 3\y(\y^T\dot\r)^2},\label{eqn:O10dotr}\\
\mathcal{O}_{20}\r %& = 
%-\dfrac{2\dot\r \dot\r^T\y  +\dot\r^T\dot\r \y}{\|\r\|^2}
%+3\dfrac{2(\y^T\dot\r) \dot\r\y^T\y +(\dot\r^T\dot\r)\y\y^T\y+(\y^T\dot\r)^2 y +2(\y^T\dot\r)\y \dot\r^T\y}{\|\r\|^2}
%- 15 \dfrac{(\y^T\dot\r)^2\y  \y^T\y}{\|\r\|^2}\nonumber \\
%&\quad -\dfrac{\ddot{\r}\y^Ty +\y^T\ddot{\r} \y +\y \ddot{\r}^T\y}{\|\r\|^1}
%+ 3 \dfrac{(\y^T\ddot{\r})\y \y^T\y}{\|\r\|}
%+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
%\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}\r,\\
%&=
%-\dfrac{2\dot\r (\dot\r^T\y)  +(\dot\r^T\dot\r) \y}{\|\r\|^2}
%+3\dfrac{2(\y^T\dot\r) \dot\r +(\dot\r^T\dot\r)\y+(\y^T\dot\r)^2 \y +2(\y^T\dot\r)^2\y }{\|\r\|^2}
%- 15 \dfrac{(\y^T\dot\r)^2\y}{\|\r\|^2}\nonumber \\
%&\quad -\dfrac{\ddot{\r} +(\y^T\ddot{\r}) \y +\y (\ddot{\r}^T\y)}{\|\r\|^1}
%+ 3 \dfrac{(\y^T\ddot{\r})\y }{\|\r\|}
%+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
%\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}\r,\\
%&=
%-\dfrac{2\dot\r (\dot\r^T\y)  +(\dot\r^T\dot\r) \y}{\|\r\|^2}
%+3\dfrac{2(\y^T\dot\r) \dot\r +(\dot\r^T\dot\r)\y+(\y^T\dot\r)^2 \y +2(\y^T\dot\r)^2\y }{\|\r\|^2}
%- 15 \dfrac{(\y^T\dot\r)^2\y}{\|\r\|^2}\nonumber \\
%&\quad -\dfrac{\ddot{\r} +(\y^T\ddot{\r}) \y +\y (\ddot{\r}^T\y)}{\|\r\|^1}
%+ 3 \dfrac{(\y^T\ddot{\r})\y }{\|\r\|}
%+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
%\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}\r,\\
%&=
%\dfrac{4\dot\r (\dot\r^T\y)  -2(\dot\r^T\dot\r)\y-6(\y^T\dot\r)^2 \y }{\|\r\|^2}
% -\mathcal{O}_{00}\ddot \r
%+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
%\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}\r,\\
& = -2\mathcal{O}_{10}\dot\r +\mathcal{O}_{00}\braces{-\ddot \r +\deriv{\ddot\r}{\r}\r},\label{eqn:O20r}\\
\mathcal{O}_{21}\r & = -2\mathcal{O}_{00} (\dot\r+\omega\times\r),\label{eqn:O21r}\\
\mathcal{O}_{21}\dot\r & = 2\mathcal{O}_{10} \dot\r
- 2 \mathcal{O}_{00}[\omega]_\times\dot\r,\label{eqn:O21rdot}
\end{align}
which are useful to derive the observability criteria.

\paragraph{Observability Rank Condition}

Now we present sufficient conditions that the observability matrix $\mathcal{O}$ has full rank.

\begin{prop}
Define three vectors $\mathbf{v}_{rel}$, $\mathbf{a}_1,\mathbf{a}_2\in\Re^3$ as 
\begin{align}
\mathbf{v}_{rel} & = \dot\r +\omega\times\r,\label{eqn:vrel}\\
\mathbf{a}_1 & = \ddot\r-\deriv{\ddot\r}{\r}\r = -2\omega\times \dot\r - [\omega]_\times^2 a \mathbf{e}_1 -\dfrac{\mu a }{\|\r_a\|^3}\mathbf{e}_1- \dfrac{3\mu\r_a^T\r}{\|\r_a\|^5}\r_a,\label{eqn:a1}\\
\mathbf{a}_2 & = \ddot\r-\deriv{\ddot\r}{\r}\r -\deriv{\ddot\r}{\dot\r}\dot\r = \mathbf{a}_1+2\omega\times\dot\r= - [\omega]_\times^2 a \mathbf{e}_1 -\dfrac{\mu a }{\|\r_a\|^3}\mathbf{e}_1- \dfrac{3\mu\r_a^T\r}{\|\r_a\|^5}\r_a.\label{eqn:a2}
\end{align}
The nonlinear relative orbital dynamics are locally weakly observable at $\x=[\r^T,\dot\r^T]$ if
%\begin{list}{}{\setlength{\leftmargin}{1cm}\setlength{\itemsep}{0mm}\setlength{\parsep}{0mm}\setlength{\topsep}{0mm}\setlength{\parskip}{0mm}\setlength{\labelwidth}{2cm}}
%\item[(i)] when $\r\times\dot\r =0$, $\r\times\mathbf{a}_1\neq 0$ and $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_1)\neq 0$,
%\item[(ii)] when $\r\times\dot\r \neq 0$, $\r\times\mathbf{a}_2\neq 0$ and $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_2)\neq 0$.
%\end{list}
\begin{alignat}{2}
(i)&\; &\text{when $\r\times\dot\r =0$, $\r\times\mathbf{v}_{ref}\neq0$, $\r\times\mathbf{a}_1\neq 0$,  and $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_1)\neq 0$},\label{eqn:cond1}\\
(ii)& &\text{when $\r\times\dot\r \neq 0$, $\r\times\mathbf{v}_{ref}\neq0$, $\r\times\mathbf{a}_2\neq 0$ and $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_2)\neq 0$.}\label{eqn:cond2}
\end{alignat}
\end{prop}

>>>>

\documentclass[letterpaper, preprint, paper,11pt]{AAS}	% for preprint proceedings
%\documentclass[letterpaper, paper,11pt]{AAS}		% for final proceedings (20-page limit)
%\documentclass[letterpaper, paper,12pt]{AAS}		% for final proceedings (20-page limit)
%\documentclass[letterpaper, paper,10pt]{AAS}		% for final proceedings (20-page limit)
%\documentclass[letterpaper, submit]{AAS}			% to submit to JAS

\usepackage{bm}
\usepackage{amssymb,amsmath,amsthm,times,graphicx,subfigure,tabularx,booktabs,colortbl,multirow,threeparttable}

\usepackage{subfigure}
%\usepackage[notref,notcite]{showkeys}  % use this to temporarily show labels
\usepackage[colorlinks=true, pdfstartview=FitV, linkcolor=black, citecolor= black, urlcolor= black]{hyperref}
\usepackage{overcite}
\usepackage{footnpag}			      	% make footnote symbols restart on each page

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\ave}[1]{\mathrm{E}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathfrak{d}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}

@@

\PaperNumber{XX-XXX}

@@

\author{Evan Kaufman\thanks{Doctoral Student, Mechanical and Aerospace Engineering, George Washington University, 801 22nd St NW, Washington, DC 20052, Email: evankaufman@gwu.edu.},  
T. Alan Lovell\thanks{Research Aerospace Engineer, Air Force Research Laboratory, Space Vehicles Directorate, Kirtland AFB, NM.},
\ and Taeyoung Lee\thanks{Assistant Professor, Mechanical and Aerospace Engineering, George Washington University, 801 22nd St NW, Washington, DC 20052, Tel: 202-994-8710, Email: tylee@gwu.edu.}
}

@@

\begin{abstract}
%An angles-only extended Kalman filter (EKF), based off of two-body motion, is developed to estimate the position of a Deputy satellite relative to a chief satellite.
A new nonlinear observability measure is proposed for relative orbit determination when lines-of-sight between satellites are measured only. It corresponds to a generalization of the observability Gramian in linear dynamic systems to the nonlinear relative orbit dynamics represented by the two-body problems. An extended Kalman filter (EKF) is adapted to this problem and is evaluated with various gravitational harmonics, initial orbital determination (IOD) predictions, and time step sizes. Preliminary results illustrate correspondence between the proposed observability measure with filtering errors. An extensive numerical analysis in realistic scenarios includes satellite propagation of the two-body problem the $J_2$ perturbation effects.
\end{abstract}

@@

Space-based surveillance or relative navigation is desirable for many spacecraft missions, such as formation control and rendezvous. Spacecraft maneuvers based only on on-board measurements reduce the total operating cost significantly, and it improves safety against communication interruptions with ground stations.  Relative navigation between spacecraft in close-proximity essentially corresponds to space-based orbit determination.  In particular, vision-based navigation and estimation of relative orbit have received attention recently, since optical sensors have desirable properties of low cost and minimal maintenance, while providing accurate line-of-sight measurements, or angles from a chief to a deputy.

@@

However, this result does not provide any information of the degree of observability. The main contribution of this paper is proposing a new  measure of observability for relative orbit determination with angles-only measurement. For a given initial condition of relative orbits, the proposed observability measure determines how much the corresponding relative orbits are easy or difficult to estimate, thereby providing a quantitative measure for relative orbit determination.

@@

% General background (very similar to the first two paragraphs of the last abstract)
%Relative navigation is desirable for many spacecraft missions, such as formation control and rendezvous, since spacecraft maneuvers based only on on-board measurements can reduce the total operating cost significantly, and it improves safety against communication interruptions with ground stations. For these close proximity missions, relative navigation essentially reduces to space-based orbit determination. Vision-based navigation and estimation problems have received attention, since optical sensors have desirable properties of having low cost and minimal maintenance, while providing accurate line-of-sight measurements.

%Relative navigation based on angles-only measurements has been the subject of much recent investigation~\cite{WofGel09,Tom11,PatLovAllRusSin12}. The problem is to determine the relative orbit between a chief spacecraft and a deputy spacecraft by using the line-of-sight measurements between the two objects, assuming that the orbit of the chief is known.

%A common problem in relative navigation with angles-only measurements is range ambiguity. Without range information, the deputy states becomes less observable. We propose a generalized observability Gramian to nonlinear systems.

%Reference 1 shows that the relative orbit is unobservable from angles-only measurements when linear relative orbital dynamics are assumed, unless there are thrusting maneuvers.
%Reference 2 investigates observability by using a relative orbit model linearized about spherical coordinates.
%Reference 3 introduces the concept of partial observability to determine a basis vector representing a family of relative orbits, and an initial orbit determination technique for this method.
%However, all of these results are based on linear relative orbital dynamics.

% Motivation: the performance of relative OD EKF is largely untested for various scenarios, including higher-fidelity gravitational models and time between measurements, for various initial predictions from IOD techniques

% Assumptions:
% 	The chief orbit initial conditions (initially circular) assumed to be well-known.
%	Measurements: a unit vector representing the direction between the chief and deputy is subject to Gaussian white noise
%		the measurements are always available (nothing blocking line-of-sight)
%		the measurements always originate from the deputy satellite
%	Initial conditions: an initial state estimate and uncertainty are assumed well-known
